PineAPPL: combining EW and QCD corrections\\ for fast evaluation of LHC processes

### 2020-10-13 paper/sec-doublecounting.tex
points 8, 10 of the report

A first example is the subtraction of (irreducible) background processes which must not be considered as such. A very blatant case
is neutral-current Drell--Yan, where the signal process is the production of an opposite-sign lepton pair, which starts
at $\mathcal O(\alpha^2)$. Because this process is usually thought
as a quark-initiated $s$-channel mechanism ($\mathrm{q} \bar{\mathrm{q}} \to \gamma^*/\mathrm{Z} \to \ell \bar{\ell}$), in many analyses the PI component,
$\gamma \gamma \to \ell \bar{\ell}$ in the $t$ channel, is considered a different process, and therefore as a background and subtracted.
The subtraction from the measured data is done by calculating the theoretical predictions of the double-photon initiated contribution, possibly including (ill-defined) higher-order
corrections. For example, in refs.~\cite{Aaboud:2017ffb,Aad:2016zzw} (a similar statement appears also in an older analysis~\cite{Aad:2013iua}), one reads:
\begin{quote}
The photon-induced process, $\gamma\gamma \to \ell \bar{\ell}$, is simulated at LO using Pythia 8
and the MRST2004qed PDF set~\cite{Martin:2004dh}. The expected yield for this process also accounts for 
NLO QED/EW corrections from references~\cite{Bardin:2012jk,Bondarenko:2013nu}, which decrease the yield by approximately \SI{30}{\percent}.
\end{quote}
Such a distinction, which is unphysical and incorrect in quantum mechanics, may be somehow justified at LO. Beyond this order, it is simply wrong.
Indeed, at $\mathcal O(\alpha^3)$, the reaction $\mathrm{q} \gamma \to \ell \bar{\ell} \mathrm{q}$ becomes possible, which
includes both kind of topologies discussed above, and needs both in order to yield an IR-finite result, see Fig.~\ref{fig:dy-pi} (as a consequence, one cannot speak of EW corrections to $\gamma \gamma \to \ell \bar{\ell}$). While it is common sense that the QCD counterpart of this subtraction should never 
be performed --- no one would
consider to \enquote{subtract} the gluon-initiated contribution to top-pair production in top analyses --- seemingly it is not so
when EW corrections are considered.

>>>>

A first example is the subtraction of (irreducible) background processes which must not be considered as such. A very blatant case
is neutral-current Drell--Yan, where the signal process is the production of an opposite-sign lepton pair, which starts
at $\mathcal O(\alpha^2)$. Because this process is usually thought
as a quark-initiated $s$-channel mechanism ($\mathrm{q} \bar{\mathrm{q}} \to \gamma^*/\mathrm{Z} \to \ell \bar{\ell}$), in many analyses the PI component,
$\gamma \gamma \to \ell \bar{\ell}$ in the $t$ channel, is considered a different process, and therefore as a background and subtracted.
The subtraction from the measured data is done by calculating the theoretical predictions of the double-photon initiated contribution, possibly including (ill-defined) higher-order
corrections. For example, in refs.~\cite{Aaboud:2017ffb,Aad:2016zzw} (a similar statement appears also in an older analysis~\cite{Aad:2013iua}), one reads:
\begin{quote}
The photon-induced process, $\gamma\gamma \to \ell \bar{\ell}$, is simulated at LO using Pythia 8
and the MRST2004qed PDF set~\cite{Martin:2004dh}. The expected yield for this process also accounts for 
NLO QED/EW corrections from references~\cite{Bardin:2012jk,Bondarenko:2013nu}, which decrease the yield by approximately \SI{30}{\percent}.
\end{quote}
Such a distinction, which is unphysical and incorrect in quantum mechanics, may be somehow justified at LO. Beyond this order, it is simply wrong.
Indeed, at $\mathcal O(\alpha^3)$, the reaction $\mathrm{q} \gamma \to \ell \bar{\ell} \mathrm{q}$ becomes possible, which
includes both kind of topologies discussed above, and needs both in order to yield an IR-finite result, see Fig.~\ref{fig:dy-pi} (as a consequence, one cannot speak of EW corrections to $\gamma \gamma \to \ell \bar{\ell}$). It may be useful for the reader to consider a QCD counterpart of this issue, 
e.g. the subtraction of gluon-initiated contribution to top-pair production, in order to get convinced of the incorrectness of this procedure.

### 2020-10-13 paper/sec-results.tex
points 8, 10 of the report

In this section we demonstrate the capabilities of the \textsc{PineAPPL}
library by interfacing it to \textsc{mg5\_aMC}, and by
computing fast-interpolation grids, accurate to NLO QCD and NLO QCD+EW,
for a common set of LHC processes in which EW corrections may sizeably
affect the accuracy of the theoretical predictions.
In order to consider realistic kinematics for these
processes, we rely on a representative set of LHC measurements. Our aim is twofold. First, we want to validate the results
obtained with \textsc{PineAPPL}; second, we want to assess the
impact of the EW corrections for the specific experimental setups. We describe
the settings employed for the computations, the corresponding results for
each process, and possible implications for the determination of PDFs.

@@

\paragraph{Experimental measurements and process features.}
We select the single-differential invariant mass distribution of the
lepton pair, $M_{\ell \bar\ell}$, measured by the ATLAS experiment at a centre-of-mass
energy of \SI{7}{\tera\electronvolt} in the high-mass region
($M_{\ell\bar\ell}>\SI{116}{\giga\electronvolt}$)~\cite{Aad:2013iua}.
We also select the single-differential rapidity distribution, $y_{\ell\bar\ell}$, in slices of
the invariant mass of the lepton pair, $M_{\ell\bar\ell}$,
measured by the CMS experiment at a centre-of-mass energy of
\SI{7}{\tera\electronvolt}~\cite{Chatrchyan:2013tia}.
These measurements are currently included as standard in the
NNPDF3.1~\cite{Ball:2017nwa} and MMHT2014~\cite{Harland-Lang:2014zoa} PDF sets,
although with appropriate kinematic cuts that remove the bins at the largest
values of invariant mass, where EW corrections become sizeable. As explained in
section~\ref{sec:pineappl-example}, 
the process has a single LO, $\mathcal{O}(\alpha^2)$; at NLO, the
QCD contribution is $\mathcal{O}(\alphas\alpha^2)$, while the EW contribution
is $\mathcal{O}(\alpha^3)$. Our NLO QCD computation includes the
$\mathcal{O}(\alpha^2)$ and $\mathcal{O}(\alphas\alpha^2)$ contributions, while
our NLO QCD+EW computation includes the $\mathcal{O}(\alpha^2)$, 
$\mathcal{O}(\alphas\alpha^2)$ and $\mathcal{O}(\alpha^3)$ contributions.
Combined QCD--EW corrections occur only at NNLO, and are therefore not
considered here. EW corrections for this process were computed in
refs.~\cite{Baur:2001ze,Arbuzov:2007db,Dittmaier:2009cr,Frederix:2018nkq} (see also~\cite{Alioli:2016fum} and references therein). The process receives contributions
from 13 (35) parton luminosities at NLO QCD (NLO QCD+EW),
see appendix~\ref{app:lumis} for details.

>>>>

In this section we demonstrate the capabilities of the \textsc{PineAPPL}
library by interfacing it to \textsc{mg5\_aMC}, and by
computing fast-interpolation grids, accurate to NLO QCD and NLO QCD+EW,
for a common set of LHC processes in which EW corrections may sizeably
affect the accuracy of the theoretical predictions.
In order to consider realistic kinematics for these
processes, we rely on a representative set of LHC measurements. Our aim is twofold. First, we want to validate the results
obtained with \textsc{PineAPPL}; second, we want to assess the
impact of the EW corrections for the specific experimental setups. We describe
the settings employed for the computations, the corresponding results for
each process, and possible implications for the determination of PDFs. It is worth to mention
that, in those phase-space regions where both NLO QCD and EW corrections are sizeable, the NNLO
correction at $\mathcal O (\alphas \alpha)$ is also expected to be important, and has either to
be included, when it is known exactly or at least in some approximate way, or to be accounted for 
as an extra uncertainty. We leave this aspect to future studies

@@

\paragraph{Experimental measurements and process features.}
We select the single-differential invariant mass distribution of the
lepton pair, $M_{\ell \bar\ell}$, measured by the ATLAS experiment at a centre-of-mass
energy of \SI{7}{\tera\electronvolt} in the high-mass region
($M_{\ell\bar\ell}>\SI{116}{\giga\electronvolt}$)~\cite{Aad:2013iua}.
We also select the single-differential rapidity distribution, $y_{\ell\bar\ell}$, in slices of
the invariant mass of the lepton pair, $M_{\ell\bar\ell}$,
measured by the CMS experiment at a centre-of-mass energy of
\SI{7}{\tera\electronvolt}~\cite{Chatrchyan:2013tia}.
These measurements are currently included as standard in the
NNPDF3.1~\cite{Ball:2017nwa} and MMHT2014~\cite{Harland-Lang:2014zoa} PDF sets,
although with appropriate kinematic cuts that remove the bins at the largest
values of invariant mass, where EW corrections become sizeable. As explained in
section~\ref{sec:pineappl-example}, 
the process has a single LO, $\mathcal{O}(\alpha^2)$; at NLO, the
QCD contribution is $\mathcal{O}(\alphas\alpha^2)$, while the EW contribution
is $\mathcal{O}(\alpha^3)$. Our NLO QCD computation includes the
$\mathcal{O}(\alpha^2)$ and $\mathcal{O}(\alphas\alpha^2)$ contributions, while
our NLO QCD+EW computation includes the $\mathcal{O}(\alpha^2)$, 
$\mathcal{O}(\alphas\alpha^2)$ and $\mathcal{O}(\alpha^3)$ contributions.
Combined QCD--EW corrections occur only at NNLO~\cite{Buccioni:2020cfi}, and are therefore not
considered here. EW corrections for this process were computed in
refs.~\cite{Baur:2001ze,Arbuzov:2007db,Dittmaier:2009cr,Frederix:2018nkq} (see also~\cite{Alioli:2016fum} and references therein). The process receives contributions
from 13 (35) parton luminosities at NLO QCD (NLO QCD+EW),
see appendix~\ref{app:lumis} for details.
